---
title: "Data Science Challenge Part 2 - NYC"
author: "Jerome Dixon"
date: "1/28/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(readxl)
library(lubridate)
library(dplyr)
library(unpivotr)
library(here)
library(data.table)
library(forecastML)


```


A senior policy official asks you to analyze why New York City and Detroit have different pedestrian crash outcomes over the last 5 years. Take Fatality Analysis Reporting System data from nhtsa.gov and compare and contrast Detroit with the New York City metropolitan area. What are the things you might analyze to understand what is different about fatal crash patterns in these cities? Explore the dataset and develop a hypothesis about why the two areas are different.

Hypothesis: Detroit pedestrian crash drivers will be younger drivers, potentially distracted with cell phone use. New York City will have less distracted drivers due to a higher operational tempo environment and how expensive it is to actually own and park a car in New York City. New York City pedestrian crash data will be driven more by volume of traffic and the number of pedestrians.

These are the key variables I believe capture or can lead to insights on this problem.

Vehicle VIN: 
Equipment Safety Features 
Vehicle HP
Vehicle Torque
Vehicle age
Vehicle Class and Model

Crash Site (GPS) and Date:
Walkability Score
Bike Lanes/Sidewalks
Pedestrian volume 
Traffic Volume
Weather patterns
Road conditions 
Traffic patterns
Intersection or Road Layout
Mandated Speed
Proximity to Local Bars/Alcohol

People:
Age of Driver
Age of Pedestrian
Alcohol/Drugs
Distracted 
Cell phone use 
Driver Origin
Driver Destination
Driver Speed
Driver Skill/Experience:
  Number of days between crash and date of 15th birthday
  Number of days between crash and date of 16th birthday 
  Number of days between crash and current birthday
  

Course of Action.

1. Research Fatality Analysis Reporting System (FARS) dataset for variables that best align to the key variables that I believe represent this problem. Dowload the last five years of history to a local directory called 'FARS'. https://www.nhtsa.gov/content/nhtsa-ftp/251

2. Once key variables from the FARS dataset have been collected, format the data into two dataframes for XGBoost. One dataframe for New York City, and one dataframe for Detroit.

3. From these two dataframes use XGBoost Feature Importance capability to see which variables drive pedestrian crash counts in each respective city.

4. Finally, compare the two XGBoost Feature Importance visualizations for insights on difference between the two cities pedestrian crash data.

(https://www.gsa.gov/reference/geographic-locator-codes-glcs-overview)
GSA GLC:
State Name	State Code	City Code	City Name	    County Code	 County Name	         	 Duty Station Code
MICHIGAN	  26	        1260    	DETROIT	      163	         WAYNE	                 261260163

NEW YORK	  36	        4170	    NEW YORK CITY	085,005,047	 RICHMOND, BRONX, KINGS	 364170085

Country Code
840



```{r Load New York Base Data 2015, warning=FALSE, message=FALSE}

# This file contains all pedestrian crash cases
pbtype_nyc_2015 <- fread("FARS/2015/PBType.csv", sep = ",", header= TRUE) %>%
  filter(str_detect(ST_CASE, "^36")) 

# Here we get number of New York Pedestrian cases
new_york_pedestrian_cases_2015 <- pbtype_nyc_2015[,2]


```


NHTSA.gov Tables of Interest: Accidents, Person, Vehicle, Distract


```{r Load New York Data 2015, warning=FALSE, message=FALSE}

# Here we get number of accidents from Pedestrian cases and then filter by New York City city code
accidents_new_york_2015 <- fread("FARS/2015/ACCIDENT.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% new_york_pedestrian_cases_2015$ST_CASE)

# Filter out New York City cases
accidents_nyc_2015 <- accidents_new_york_2015 %>%  filter(CITY == '4170')

#City of New York Pedestrian Cases
nyc_pedestrian_cases_2015 <- as.data.frame(accidents_nyc_2015$ST_CASE)
names(nyc_pedestrian_cases_2015) <- "ST_CASE"

# Add Date Column
accidents_nyc_2015$Date <- paste0(accidents_nyc_2015$YEAR,"-",accidents_nyc_2015$MONTH,"-", accidents_nyc_2015$DAY)
accidents_nyc_2015$Date <- as.Date(accidents_nyc_2015$Date)

accidents_nyc_2015 %<>% select(53,1:52)

# Fill in gaps in Time Series
accidents_nyc_2015 %<>% fill_gaps(., date_col = 1, frequency = '1 day', groups = NULL, static_features = NULL)


```


```{r Add Data to 2015 Dataframe, warning=FALSE, message=FALSE}


person_nyc_2015 <- fread("FARS/2015/person.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE) 

accidents_nyc_2015 <- left_join(accidents_nyc_2015, person_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))
  
  

distractions_nyc_2015 <- fread("FARS/2015/Distract.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE)

accidents_nyc_2015 <- left_join(accidents_nyc_2015, distractions_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


driver_impair_nyc_2015 <- fread("FARS/2015/DrImpair.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE)

accidents_nyc_2015 <- left_join(accidents_nyc_2015, driver_impair_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmcrash_nyc_2015 <- fread("FARS/2015/nmcrash.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE)

accidents_nyc_2015 <- left_join(accidents_nyc_2015, nmcrash_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmimpair_nyc_2015 <- fread("FARS/2015/NMImpair.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE)

accidents_nyc_2015 <- left_join(accidents_nyc_2015, nmimpair_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmprior_nyc_2015 <- fread("FARS/2015/NMPrior.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2015$ST_CASE)

accidents_nyc_2015 <- left_join(accidents_nyc_2015, nmprior_nyc_2015, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))

accidents_nyc_2015[is.na(accidents_nyc_2015)] <- 0


```



```{r View Column Names to Prune Unnecessary Variables}

names_cleanup <- names(accidents_nyc_2015)

names_cleanup

```



# Again we select the variables that align best for our initial hypothesis.

```{r Final 2015 Dataframe}

accidents_nyc_2015 %<>% select("Date", "ST_CASE","DAY_original","MONTH_original",   "DAY_WEEK","HOUR_original","MINUTE_original","FUNC_SYS_original","ROUTE","TWAY_ID","TWAY_ID2",  "LATITUDE","LONGITUD","TYP_INT","PER_NO_original","MAKE","MAK_MOD","BODY_TYP","MOD_YEAR","AGE","SEX","DRINKING","DRUGS","DRUG_DET","RACE","LOCATION","DRIMPAIR") 

accidents_nyc_2015 %<>% unique()

#write.csv(accidents_nyc_2015, "nyc-2015.csv")


```



```{r Load New York Data 2016, warning=FALSE, message=FALSE}

# This file contains all pedestrian crash cases
pbtype_nyc_2016 <- fread("FARS/2016/PBType.csv", sep = ",", header= TRUE) %>%
  filter(str_detect(ST_CASE, "^36")) 

# Here we get number of New York Pedestrian cases
new_york_pedestrian_cases_2016 <- as.data.frame(pbtype_nyc_2016$ST_CASE)
names(new_york_pedestrian_cases_2016) <- "ST_CASE"


# Here we get number of accidents from Pedestrian cases and then filter by New York City city code
accidents_nyc_2016 <- fread("FARS/2016/ACCIDENT.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% new_york_pedestrian_cases_2016$ST_CASE)

# Filter out nyc City cases
accidents_nyc_2016 %<>%  filter(CITY == '4170')

#City of nyc Pedestrian Cases
nyc_pedestrian_cases_2016 <- as.data.frame(accidents_nyc_2016$ST_CASE)
names(nyc_pedestrian_cases_2016) <- "ST_CASE"


# Add Date Column
accidents_nyc_2016$Date <- paste0(accidents_nyc_2016$YEAR,"-",accidents_nyc_2016$MONTH,"-", accidents_nyc_2016$DAY)

# Format Date column for fill_gap function
accidents_nyc_2016$Date <- as.Date(accidents_nyc_2016$Date)
accidents_nyc_2016 %<>% select(92,3,1,2,4:91)


# Fill in gaps in Time Series
accidents_nyc_2016 %<>% fill_gaps(., date_col = 1, frequency = '1 day', groups = NULL, static_features = NULL)


```



```{r Add Data to 2016 Dataframe}


person_nyc_2016 <- fread("FARS/2016/person.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE) 

accidents_nyc_2016 <- left_join(accidents_nyc_2016, person_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))
  
  

distractions_nyc_2016 <- fread("FARS/2016/Distract.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE)

accidents_nyc_2016 <- left_join(accidents_nyc_2016, distractions_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


driver_impair_nyc_2016 <- fread("FARS/2016/DrImpair.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE)

accidents_nyc_2016 <- left_join(accidents_nyc_2016, driver_impair_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmcrash_nyc_2016 <- fread("FARS/2016/nmcrash.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE)

accidents_nyc_2016 <- left_join(accidents_nyc_2016, nmcrash_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmimpair_nyc_2016 <- fread("FARS/2016/NMImpair.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE)

accidents_nyc_2016 <- left_join(accidents_nyc_2016, nmimpair_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmprior_nyc_2016 <- fread("FARS/2016/NMPrior.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2016$ST_CASE)

accidents_nyc_2016 <- left_join(accidents_nyc_2016, nmprior_nyc_2016, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))

accidents_nyc_2016[is.na(accidents_nyc_2016)] <- 0


```


```{r View 2016 Column Names for Pruning}


names_cleanup <- names(accidents_nyc_2016)

names_cleanup


```


```{r Final 2016 Dataframe}

accidents_nyc_2016 %<>% select("Date", "ST_CASE","DAY_original","MONTH_original",   "DAY_WEEK","HOUR_original","MINUTE_original","FUNC_SYS_original","ROUTE","TWAY_ID","TWAY_ID2",  "LATITUDE","LONGITUD","TYP_INT","PER_NO_original","MAKE","MAK_MOD","BODY_TYP","MOD_YEAR","AGE","SEX","DRINKING","DRUGS","DRUG_DET","RACE","LOCATION","DRIMPAIR") 

accidents_nyc_2016 %<>% unique()
#write.csv(accidents_nyc_2016, "nyc-2016.csv")


```



```{r Load New York Data 2017, warning=FALSE, message=FALSE}


# This file contains all pedestrian crash cases
pbtype_nyc_2017 <- fread("FARS/2017/PBType.csv", sep = ",", header= TRUE) %>%
  filter(str_detect(ST_CASE, "^36")) 

# Here we get number of new_york Pedestrian cases
new_york_pedestrian_cases_2017 <- as.data.frame(pbtype_nyc_2017$ST_CASE)
names(new_york_pedestrian_cases_2017) <- "ST_CASE"


# Here we get number of accidents from Pedestrian cases and then filter by nyc city code
accidents_nyc_2017 <- fread("FARS/2017/ACCIDENT.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% new_york_pedestrian_cases_2017$ST_CASE)

# Filter out nyc City cases
accidents_nyc_2017 %<>%  filter(CITY == '4170')

#City of nyc Pedestrian Cases
nyc_pedestrian_cases_2017 <- as.data.frame(accidents_nyc_2017$ST_CASE)
names(nyc_pedestrian_cases_2017) <- "ST_CASE"

accidents_nyc_2017$Date <- paste0(accidents_nyc_2017$YEAR,"-",accidents_nyc_2017$MONTH,"-", accidents_nyc_2017$DAY)

# Add Date Column
accidents_nyc_2017$Date <- as.Date(accidents_nyc_2017$Date)

accidents_nyc_2017 %<>% select(92,3,1,2,4:91)

# Fill in gaps in Time Series
accidents_nyc_2017 %<>% fill_gaps(., date_col = 1, frequency = '1 day', groups = NULL, static_features = NULL)



```


```{r Add Data to 2017 Dataframe}


# Here we get the number of people attached to each Pedestrian case
person_nyc_2017 <- fread("FARS/2017/person.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE) 

accidents_nyc_2017 <- left_join(accidents_nyc_2017, person_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))
  
  

distractions_nyc_2017 <- fread("FARS/2017/Distract.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE)

accidents_nyc_2017 <- left_join(accidents_nyc_2017, distractions_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


driver_impair_nyc_2017 <- fread("FARS/2017/DrImpair.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE)

accidents_nyc_2017 <- left_join(accidents_nyc_2017, driver_impair_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmcrash_nyc_2017 <- fread("FARS/2017/nmcrash.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE)

accidents_nyc_2017 <- left_join(accidents_nyc_2017, nmcrash_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmimpair_nyc_2017 <- fread("FARS/2017/NMImpair.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE)

accidents_nyc_2017 <- left_join(accidents_nyc_2017, nmimpair_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmprior_nyc_2017 <- fread("FARS/2017/NMPrior.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2017$ST_CASE)

accidents_nyc_2017 <- left_join(accidents_nyc_2017, nmprior_nyc_2017, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))

accidents_nyc_2017[is.na(accidents_nyc_2017)] <- 0


```



```{r View 2017 Column Names for Pruning}

names_cleanup <- names(accidents_nyc_2017)

names_cleanup


```


```{r Final 2017 Dataframe}

accidents_nyc_2017 %<>% select("Date", "ST_CASE","DAY_original","MONTH_original",   "DAY_WEEK","HOUR_original","MINUTE_original","FUNC_SYS_original","ROUTE","TWAY_ID","TWAY_ID2",  "LATITUDE","LONGITUD","TYP_INT","PER_NO_original","MAKE","MAK_MOD","BODY_TYP","MOD_YEAR","AGE","SEX","DRINKING","DRUGS","DRUG_DET","RACE","LOCATION","DRIMPAIR") 

accidents_nyc_2017 %<>% unique()
#write.csv(accidents_nyc_2017, "nyc-2017.csv")


```



```{r Load New York Data 2018, warning=FALSE, message=FALSE}

# This file contains all pedestrian crash cases
pbtype_nyc_2018 <- fread("FARS/2018/PBType.csv", sep = ",", header= TRUE) %>%
  filter(str_detect(ST_CASE, "^36")) 

# Here we get number of new_york Pedestrian cases
new_york_pedestrian_cases_2018 <- as.data.frame(pbtype_nyc_2018$ST_CASE)
names(new_york_pedestrian_cases_2018) <- "ST_CASE"


# Here we get number of accidents from Pedestrian cases and then filter by nyc city code
accidents_nyc_2018 <- fread("FARS/2018/ACCIDENT.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% new_york_pedestrian_cases_2018$ST_CASE)

# Filter out nyc City cases
accidents_nyc_2018 %<>%  filter(CITY == '4170')

#City of nyc Pedestrian Cases
nyc_pedestrian_cases_2018 <- as.data.frame(accidents_nyc_2018$ST_CASE)
names(nyc_pedestrian_cases_2018) <- "ST_CASE"


# Add Date Column
accidents_nyc_2018$Date <- paste0(accidents_nyc_2018$YEAR,"-",accidents_nyc_2018$MONTH,"-", accidents_nyc_2018$DAY)
accidents_nyc_2018$Date <- as.Date(accidents_nyc_2018$Date)

accidents_nyc_2018 %<>% select(92,2,3,1,4:91)

# Fill in gaps in Time Series
accidents_nyc_2018 %<>% fill_gaps(., date_col = 1, frequency = '1 day', groups = NULL, static_features = NULL)


```



```{r Add Data to 2018 Dataframe}

# Here we get the number of people attached to each Pedestrian case
person_nyc_2018 <- fread("FARS/2018/person.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE) 

accidents_nyc_2018 <- left_join(accidents_nyc_2018, person_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))
  
  

distractions_nyc_2018 <- fread("FARS/2018/Distract.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE)

accidents_nyc_2018 <- left_join(accidents_nyc_2018, distractions_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


driver_impair_nyc_2018 <- fread("FARS/2018/DrImpair.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE)

accidents_nyc_2018 <- left_join(accidents_nyc_2018, driver_impair_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmcrash_nyc_2018 <- fread("FARS/2018/nmcrash.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE)

accidents_nyc_2018 <- left_join(accidents_nyc_2018, nmcrash_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmimpair_nyc_2018 <- fread("FARS/2018/NMImpair.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE)

accidents_nyc_2018 <- left_join(accidents_nyc_2018, nmimpair_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmprior_nyc_2018 <- fread("FARS/2018/NMPrior.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2018$ST_CASE)

accidents_nyc_2018 <- left_join(accidents_nyc_2018, nmprior_nyc_2018, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))

accidents_nyc_2018[is.na(accidents_nyc_2018)] <- 0



```



```{r View 2018 Column Names for Pruning}


names_cleanup <- names(accidents_nyc_2018)

names_cleanup

```




```{r Final 2018 Dataframe}

accidents_nyc_2018 %<>% select("Date", "ST_CASE","DAY_original","MONTH_original",   "DAY_WEEK","HOUR_original","MINUTE_original","FUNC_SYS_original","ROUTE","TWAY_ID","TWAY_ID2",  "LATITUDE","LONGITUD","TYP_INT","PER_NO_original","MAKE","MAK_MOD","BODY_TYP","MOD_YEAR","AGE","SEX","DRINKING","DRUGS","DRUG_DET","RACE","LOCATION","DRIMPAIR") 

accidents_nyc_2018 %<>% unique()


```



```{r Load New York Data 2019, warning=FALSE, message=FALSE}


# This file contains all pedestrian crash cases
pbtype_nyc_2019 <- fread("FARS/2019/PBType.csv", sep = ",", header= TRUE) %>%
  filter(str_detect(ST_CASE, "^36")) 

# Here we get number of new_york Pedestrian cases
new_york_pedestrian_cases_2019 <- as.data.frame(pbtype_nyc_2019$ST_CASE)
names(new_york_pedestrian_cases_2019) <- "ST_CASE"


# Here we get number of accidents from Pedestrian cases and then filter by nyc city code
accidents_nyc_2019 <- fread("FARS/2019/ACCIDENT.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% new_york_pedestrian_cases_2019$ST_CASE)

# Filter out nyc City cases
accidents_nyc_2019 %<>%  filter(CITY == '4170')

#City of nyc Pedestrian Cases
nyc_pedestrian_cases_2019 <- as.data.frame(accidents_nyc_2019$ST_CASE)
names(nyc_pedestrian_cases_2019) <- "ST_CASE"


# Add Date Column
accidents_nyc_2019$Date <- paste0(accidents_nyc_2019$YEAR,"-", accidents_nyc_2019$MONTH,"-", accidents_nyc_2019$DAY)

accidents_nyc_2019$Date <- as.Date(accidents_nyc_2019$Date)

accidents_nyc_2019 %<>% select(92,3,1,2,4:91)

# Fill in gaps in Time Series
accidents_nyc_2019 %<>% fill_gaps(., date_col = 1, frequency = '1 day', groups = NULL, static_features = NULL)


```



```{r Add Data to 2019 Dataframe}


# Here we get the number of people attached to each Pedestrian case
person_nyc_2019 <- fread("FARS/2019/person.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE) 

accidents_nyc_2019 <- left_join(accidents_nyc_2019, person_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))
  
  
distractions_nyc_2019 <- fread("FARS/2019/Distract.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE)

accidents_nyc_2019 <- left_join(accidents_nyc_2019, distractions_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


driver_impair_nyc_2019 <- fread("FARS/2019/DrImpair.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE)

accidents_nyc_2019 <- left_join(accidents_nyc_2019, driver_impair_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmcrash_nyc_2019 <- fread("FARS/2019/nmcrash.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE)

accidents_nyc_2019 <- left_join(accidents_nyc_2019, nmcrash_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmimpair_nyc_2019 <- fread("FARS/2019/NMImpair.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE)

accidents_nyc_2019 <- left_join(accidents_nyc_2019, nmimpair_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


nmprior_nyc_2019 <- fread("FARS/2019/NMPrior.csv", sep = ",", header= TRUE) %>%
	filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE)

accidents_nyc_2019 <- left_join(accidents_nyc_2019, nmprior_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


race_nyc_2019 <- fread("FARS/2019/Race.csv", sep = ",", header= TRUE) %>%
  filter(ST_CASE %in% nyc_pedestrian_cases_2019$ST_CASE) 

accidents_nyc_2019 <- left_join(accidents_nyc_2019, race_nyc_2019, by=c("ST_CASE"),
                                    suffix = c("_original", "_new"))


```


```{r View 2019 Column Names for Pruning}


names_cleanup <- names(accidents_nyc_2019)

names_cleanup

```


```{r Final 2019 Dataframe}

names(accidents_nyc_2019)[118] <- "MAKE"
names(accidents_nyc_2019)[120] <- "MAK_MOD"
names(accidents_nyc_2019)[121] <- "BODY_TYP"
names(accidents_nyc_2019)[123] <- "MOD_YEAR"
names(accidents_nyc_2019)[137] <- "AGE"
names(accidents_nyc_2019)[139] <- "SEX"
names(accidents_nyc_2019)[159] <- "DRINKING"
names(accidents_nyc_2019)[169] <- "DRUGS"
names(accidents_nyc_2019)[171] <- "DRUG_DET"
names(accidents_nyc_2019)[219] <- "DRIMPAIR"


accidents_nyc_2019 %<>% select("Date", "ST_CASE","DAY_original","MONTH_original",   "DAY_WEEK","HOUR_original","MINUTE_original","FUNC_SYS_original","ROUTE","TWAY_ID","TWAY_ID2",  "LATITUDE","LONGITUD","TYP_INT","PER_NO_original","MAKE","MAK_MOD","BODY_TYP","MOD_YEAR","AGE","SEX","DRINKING","DRUGS","DRUG_DET","RACE","LOCATION","DRIMPAIR") 

accidents_nyc_2019 %<>% unique()


```



```{r Merge Dataframes for XGBoost}


# Merge all files and remove duplicate rows
nyc_pedestrians <- rbind(accidents_nyc_2015, accidents_nyc_2016, accidents_nyc_2017, accidents_nyc_2018, accidents_nyc_2019) 


```


```{r Save Final Dataframe for Future Analysis}

write.csv(nyc_pedestrians, "nyc-pedestrians.csv")

```

